TSELM: Target speaker extraction using discrete tokens and language models

### 2024-09-09 src/paper.tex
update

In contrast to blind speech separation, which seeks to isolate individual utterances from a mixture of known speakers, target speaker extraction (TSE) focuses on extracting only the voice of the target speaker using auxiliary information specific to that speaker. Current models are predominantly discriminative, employing masking strategies to directly minimize the distance between the estimated and clean speech signals \cite{luo2019conv,spex_plus,sepformer,sef_net}. However, these discriminative approaches often struggle to generalize to unseen data and may introduce undesirable distortions \cite{distortion}. To address these limitations, generative models have been proposed. These models aim to learn the underlying distribution of the target speech, rather than mapping mixed speech to clean speech. Generative models have been shown to achieve performance comparable to that of discriminative models \cite{target_diff,tokensplit} in blind speech separation and TSE.

>>>>

In contrast to blind speech separation, which seeks to isolate individual utterances from a mixture of known speakers, target speaker extraction (TSE) focuses on extracting only the voice of the target speaker using auxiliary information. Current models are predominantly discriminative, employing masking strategies to directly minimize the distance between the estimated and clean speech signals \cite{luo2019conv,spex_plus,sepformer,sef_net}. However, these discriminative approaches often struggle to generalize to unseen data and may introduce undesirable distortions \cite{distortion}. To address these limitations, generative models have been proposed. These models aim to learn the underlying distribution of the target speech, rather than mapping mixed speech to clean speech. Generative models have been shown to achieve performance comparable to that of discriminative models \cite{target_diff,tokensplit} in blind speech separation and TSE.
